---
title: "R_commands"
author: "Tiago Nardi"
date: "23/3/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## PACKAGES
Install and load a package
``` {r, eval=FALSE}
install.packages("package_name")
library(package_name)
```
## VARIABLES AND VECTORS
Create a variable
``` {r, eval=FALSE}
variable_name <- "object_name"
```
Create a vector
``` {r, eval=FALSE}
vector_name <- c("object_1", "object_2", "object_3", …)
```
## DATAFRAMES
Read a file
``` {r, eval=FALSE}
read.csv("File_name", header = T, sep="symbol", quote = "symbol",
         dec = "symbol", na.strings = value/values)
read.table("File_name", header = T, sep = "symbol",
           quote = "symbol", dec = "symbol", na.strings = value/values)
# requires to load the package "readxl"
read_excel("File_name", sheet = number, na = value/values)
```
Write a file
``` {r, eval=FALSE}
write.csv(dataframe_name, file="Output_name.csv", row.names=F)
```
To compactly display the structure of a dataframe
``` {r, eval=FALSE}
head(dataframe_name)
str(dataframe_name)
summary(dataframe_name)
```
Get row names and column names
``` {r, eval=FALSE}
rownames(dataframe_name)
colnames(dataframe_name)
```
Check the object type
``` {r, eval=FALSE}
class(object_name)
```
Change the object type
``` {r, eval=FALSE}
as.character(dataframe_name$column_name)
as.factor(dataframe_name$column_name)
as.numeric(dataframe_name$column_name)
as.integer(dataframe_name$column_name)
as.data.frame(matrix_name)
```
Check if a value is present in a column
``` {r, eval=FALSE}
"value" %in% dataframe_name$column_name
```
Get an object from a column:
``` {r, eval=FALSE}

dataframe_name$column_name[position in the column]
```
Extract the first location that has the value of interest
``` {r, eval=FALSE}
match("value", dataframe_name$column_name)
```
Extract all locations that have the value of interest
``` {r, eval=FALSE}
which(dataframe_name$column_name %in% "value")
```
Modify an object from a column
``` {r, eval=FALSE}
dataframe_name$column_name[position in the column] <- "new_value"
```
Extract objects from a dataframe
``` {r, eval=FALSE}
Specific object:dataframe_name["row_name", "column_name"]
dataframe_name[row_index, column_index]
```
Single column
``` {r, eval=FALSE}
dataframe_name$column_name
dataframe_name[ , "column_name"]
dataframe_name[ , column_index]
```

Single row
``` {r, eval=FALSE}
dataframe_name["row_name", ]
dataframe_name[row_index, ]
```
Get more columns
``` {r, eval=FALSE}
dataframe_name[ , c("column_1_name", "column_2_name")]
dataframe_name[ , c(first_column_index, second_column_index)]
```
Get more rows
``` {r, eval=FALSE}
dataframe_name[c("row_1_name", "row_2_name"), ]
dataframe_name[c(first_row_index, second_row_index), ]
```

Get a range of columns
``` {r, eval=FALSE}
dataframe_name[ , c(first_column_index : last_column_index)]
```
Get a range of rows
``` {r, eval=FALSE}
dataframe_name[c(first_row_index : last_row_index), ]
```
Add a new column
``` {r, eval=FALSE}
dataframe_name$column_name <- new_column
cbind(dataframe_name, new_column)
cbind(new_column, dataframe_name)
cbind(dataframe_name[range_index_previous_column], new_column_name,
dataframe_name[range_indices_following_columns])
```
Remove a column
``` {r, eval=FALSE}
dataframe_name$column_name <- NULL
```
Combine two columns into a new one
``` {r, eval=FALSE}
paste(dataframe_name$column_name_1, dataframe_name$column_name_2, sep="value")
```
Separate one column into two columns
``` {r, eval=FALSE}
library(tidyr)
separate(dataframe_name, column_name, into=c("column_name_1", "column_name_2"), sep= "value")
```
Get the largest element of a column
``` {r, eval=FALSE}
max(dataframe_name$column_name)
```
Get the smallest element of a column
``` {r, eval=FALSE}
min(dataframe_name$column_name)
```
Get minimum and maximum of all the given arguments:
``` {r, eval=FALSE}
range(dataframe_name$column_name)
```
Get the length of a column
``` {r, eval=FALSE}
length(vector_name)
```
Get a vector of unique values
``` {r, eval=FALSE}
unique(dataframe_name$column_name)
```
Sort the values in a column
``` {r, eval=FALSE}
sort(dataframe_name$column_name)
```
Combine two data frames
``` {r, eval=FALSE}
rbind(dataframe_1_name, dataframe_2_name)
```
Rename a column
``` {r, eval=FALSE}
colnames(dataframe_name)[column_index] <- "new_column_name"
```
Create a new dataframe from another one
``` {r, eval=FALSE}
cbind(dataframe_name$column_1, dataframe_name$column_2, dataframe_name$column_3)
```
Merge two dataframes on the basis of a column
``` {r, eval=FALSE}
new_dataframe_name <- merge(dataframe_1_name, dataframe_2_name, by = column_name)
```
Subset
``` {r, eval=FALSE}
subset(dataframe_name, dataframe_name$column logical_operator logical expression)
subset(dataframe_name, dataframe_name$column logical_operator logical expression &
dataframe_name$column2 logical_operator logical expression2)
subset(dataframe_name, dataframe_name$column logical_operator logical expression |
dataframe_name$column2 logical_operator logical expression2)
```
Drop levels:
``` {r, eval=FALSE}
subset_dataframe_name <- droplevels(subset_dataframe_name)
```
Build a contingency table
``` {r, eval=FALSE}
table(dataframe_name$column_1, dataframe_name$column_2)
```
## GRAPHS
Visualize a distribution:
``` {r, eval=FALSE}
hist(dataframe_name$column, breaks = "value", col = "colour_name", border = "color_name", main = "Title",
xlab = "x_name", ylab = "y_name")
- Visualize a distribution of grouped values:
barplot(table, col = "colour_name", beside = TRUE, cex.names = number, las = number, main = "Title", xlab =
"x_name", ylab = "y_name")
- Add the legend to a plot:
legend(coordinates, legend_text, fill = "colour_name", col = "colour_name", bg = "colour_name", pch=number)
legend("topright", legend_text, col = "colour_name", pch = 15)
```
Save a graph:
``` {r, eval=FALSE}
pdf("Graph_name.pdf")
graph_command
legend_command
dev.off()
```
## STATISTICAL ANALYSIS
``` {r, eval=FALSE}
shapiro.test(numeric_vector_data_values)
wilcox.test(numeric_vector_data_values ~ binary_vector_data_values)
table_chi_square <- table(dataframe_name$column_1, dataframe_name$column_2)
chisq.test(table_chi_square)
```
## HELP
``` {r, eval=FALSE}
?function_name
help(function_name)
```